---
title: "MassFluxes_GSA2020"
author: "William Nguyen"
date: "10/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

ALRIGHT! Let's derive some dang mass fluxes. First, reading in the appropriate packages and csv file for the chemistry data.

```{r}
library(tidyverse) # godsend: ggplot2, tibble, tidyr, readr, dplyr, purrr, stringr, forcats
library(dplyr)

df <- read.csv("https://raw.githubusercontent.com/wdnguyen/CR/master/CostaRica_Chemistry_20200920.csv", na.strings = "BDL", stringsAsFactors = FALSE)
# could also use df[df == "BDL"] <- NA

df$SamplingDate <- as.POSIXct(df$SamplingDate, format = "%m/%d/%Y %H:%M")
df$Chemetrics_Acidified_Date <- as.POSIXct(df$Chemetrics_Acidified_Date, format = "%m/%d/%Y %H:%M")

# Let's just focus on the 2018 data
df <- df[format(df$SamplingDate, '%Y') != "2019", ]

# Let's also convert all negative concentrations to NaN
df <- df %>%
  mutate_all(funs(replace(., .<0, NA)))
```

Assigning colors to endmembers and mixes:

```{r RGB colors}
rainc <- rgb(0.63,0.32,0.18) # sienna
soilc <- rgb(0.11,0.58,0.08) # green
springc <- rgb(1,0.71,0.1) # yellow
usc <- rgb(0.67,0.06,0.06) # red
dsc <- rgb(0.06,0.22,0.67) # blue
```

I'm going to quickly attempt a last-desperate attempt at EMMA (pca :( right now).

First, we need to figure out what to do with the missing data. Median replacement? It would be good if I figured out how to replace by endmember group. For now, I will be using a function devloped by someone else:

```{r Installing missing value function}
# source("dineof.R")
# http://menugget.blogspot.com/2012/10/dineof-data-interpolating-empirical.html
# JK didn't work
```

```{r defining pca}
# df <- dineof(df)
df <- df %>% 
   mutate_all(~ifelse(is.na(.), median(., na.rm = TRUE), .))
df.pca <- prcomp(df[,c(20:27)], center = TRUE, scale. = TRUE) # let's just try with the base cations, nutrients, and select few elements
summary(df.pca)
```

