---
title: "MissingData"
author: "William Nguyen"
date: "10/4/2020"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Imputing missing water quality data in the Howler Monkey Watershed

There's a couple ad-hoc ways of dealing with miss data, such as listwise deletion and mean imputation. Listwise deletion only really works when you have a few missing values and mean imputation brings too much bias to variances and covariances. In this example, I'll be using Amelia to perform multiple imputation, which should give me some probable predictions for my missing values.

In the future, I may want to implement some more robust means of imputation, such as CART, random forest, and artificial neural networks, but I'll save that for the manuscript. 

```{r Installing libraries, include = FALSE, warning=FALSE}
library(Amelia)
library(tidyverse)
library(dplyr)
library(ggthemes)
library(naniar) # for replace functions
library(VIM)
library(FactoMineR)
library(missMDA)
```

OK first, I'm going to take a stab as missMDA in R.
```{r}
#library(missMDA)
data("orange")
nb <- estim_ncpPCA(orange, scale=TRUE) ## Estimate no. of dimensions
comp <- imputePCA(orange, ncp=2, scale=TRUE) ## Impute the table
res.pca <- PCA(comp$completeObs) ## Do the PCA
```

Visualizing uncertainty due to missing data with missMDA:
```{r}
mi <- MIPCA(orange, scale = TRUE, ncp = 2)
plot(mi)
```

Another demo using genotype-environment data set (from vignette)

```{r missMDA on geno data}
set.seed(1234)
# library(missMDA)
data("geno", package = "missMDA")
summary(geno)
head(round(geno,2))
```
```{r missMDA on geno data pt 2}
ncomp <- estim_ncpPCA(geno, ncp.min=0, ncp.max=6) # by default, estim_ncpPCA using GCV method, could using Kfold by specifying method.cv="Kfold" (would be good to add two individual argumnts, pNA = 0.05, nbsim=100)
ncomp$ncp # returns NSEP for no. of dimensions minimizing the MSEP in object ncp (should be 2-D)
res.imp <- imputePCA(geno, ncp = 2, scale = TRUE, method = "Regularized", row.w = NULL, coeff.ridge = 1, threshold = 1e-06, seed = NULL, nb.init = 1, maxiter = 1000)
head(round(res.imp$completeObs, 2))
res.pca <- PCA(res.imp$completeObs)
```

To perform multiple imputation, we use the MIPCA function, which generates multiple imputed data sets. Then, the plot method for 'MIPCA' objects is applied to visualize the impact of the different imputed values on the PCA results.

```{r Multiple imputation for missMDA}
resMIPCA <- MIPCA(geno, ncp = 2, nboot = 200)
plot(resMIPCA)
```

```{r MIPCA but with the arguments}
plot(resMIPCA, choice = "all", axes = c(1,2), new.plot = TRUE, main = NULL, level.conf = 0.95)
```


```{r Reading in data}
df <- read.csv("https://raw.githubusercontent.com/wdnguyen/CR/master/CostaRica_Chemistry_Q_20201002.csv", stringsAsFactors = FALSE)
df$SamplingDate <- as.POSIXct(df$SamplingDate, format ="%m/%d/%Y %H:%M")

### For now, let's set BDL as 0.1, globally (across all data)
df <- dplyr::mutate_if(tibble::as_tibble(df), # changed as.tibble to as_tibble
                       is.character,
                       stringr::str_replace_all, pattern = "BDL", replacement = "0.1")

### Convert character vectors into numeric (hopefully blanks become NA)
df[,c(12,13,15,17,19,24,28)] <- sapply(df[,c(12,13,15,17,19,24,28)], as.numeric) # NH3_f, H2S_f, Fetot, Mn_f, As_f, F, Li

### And for good measure, I will replace all negative concentrations with NA. https://cran.r-project.org/web/packages/naniar/vignettes/replace-with-na.html
df <- df %>%
  replace_with_na_at(.vars = c("Mo","U","Al","Si","Ca","Cr","Fe","Ti","B","Ba","As"),
                     condition = ~.x < 0) # I could have just used replace_with_na_all, but O18 and D have natural negative values

# For some reason, df ID is coming out all weird
names(df)[1] <- "ID"

# Let's also just drop some columns that are REALLY lacking data: AlkalinityS, Mn_f, SO4_f, As_f
df <- subset(df, select = -c(AlkalinityS, O2, SO4_f, As_f, B, Ba, Fetot, Mn_f, TotalNField, Sr2, U, Mo, Cd, Sb, Pb))


# Remove DS9Bot, Howler Weir
df <- df[-c(21, 31),] 

### 2018 df
df2018 <- df[format(df$SamplingDate, '%Y') != "2019", ]

### Subsetting data by site / end-member
DS <- subset(df, Site == "Downstream")
US <- subset(df, Site == "Upstream")
SOIL <- subset(df, Site == "Soil")
RAIN <- subset(df, Site == "Rain")
SPRING <- subset(df, Site == "Spring")

### Just 2018 Data
DS2018 <- DS[format(DS$SamplingDate, '%Y') != "2019", ]
US2018 <- US[format(US$SamplingDate, '%Y') != "2019", ]
SOIL2018 <- SOIL[format(SOIL$SamplingDate, '%Y') != "2019", ]
RAIN2018 <- RAIN[format(RAIN$SamplingDate, '%Y') != "2019", ]
SPRING2018 <- SPRING[format(SPRING$SamplingDate, '%Y') != "2019", ]
```

OK cool, let's do a summary of df:

```{r summary}
summary(df)
```

So, for my imputations, it probably makes sense to do them by end member/group.

First, I'll do a rough PCA with not-so-great imputation. The incomplete data set can be imputed using the function imputePCA performing the iterative PCA algorithm, specifying the number of dimensions through the argument ncp=2. At convergence the algorithm provides both an estimation of the scores and loadings as well as a completed data set. The imputePCA function outputs the imputed data set. The completed data set is in the object completeObs. The imputePCA function also outputs the fitted matrix X^ in the object fitted. http://juliejosse.com/wp-content/uploads/2018/05/DataAnalysisMissingR.html

```{r PCA with missing values using imputePCA}
### must have missMDA
# df1 <- df[,c(6:9,11:14,20:60)]
# np <- estim_ncpPCA(df1, method.cv = "Kfold", verbose = FALSE)
# plot(0:5, np$criterion, xlab = "np dim", ylab = "MSEP")
# res.comp <- imputePCA(df1, ncp = np$ncp) # iterativePCA algorithm
# res.comp$completeObs[1:3,]
# imp <- cbind.data.frame(res.comp$completeObs,df$Site)
# res.pca <- PCA(imp, quanti.sup = 12, quali.sup = 50, ncp = np$ncp, graph=FALSE) ### Cl (12) as quan, Site (50) as qual
# plot(res.pca, hab=50, lab="quali");
# plot(res.pca, choix="var")
```

OK, now for the real deal with Amelia and multiple imputation, which assumes Joint Modeling (one joint probabilistic model for the variables all together)

```{r Amelia}
# ignore_vars <- c("ID", "Chemetrics_Acidified_Date", "Source", "Q", "Seconds")
# a.out <- amelia(df,                   # data set
#                m = 5,                # number of imputations, usually 5
#                ts = "SamplingDate",  # time series
#                cs = "Site",          # cross section
#                idvars = ignore_vars) # vars to leave out of imputation
```

```{r}
#ignore_vars <- c("ID", "Site", "Chemetrics_Acidified_Date", "Source", "Q", "Seconds", "Temp", "pH", "ORP")
#a.out <- amelia(DS2019,                   # data set
#                m = 5,                # number of imputations, usually 5
#                ts = "SamplingDate",  # time series
#                idvars = ignore_vars, # vars to leave out of imputation
#                empri = 20)
```

OK, so Amelia flopped hard. I have too many missing values to imputate. So let's try a different approach. 

```{r Let's impute by end member group / mixing sample}
## HM.....let's just take out the variables that matter, for example in DS2018
ds1 <- data.frame(DS2018) # Take a copy of the data

# let's also set ID as the rowname
rownames(ds1) <- ds1$ID

# Let's try looking at all good enough variables. First, I'll use a function to help determine the value of dimension S that that minimizes the mean squared error of prediction (MSEP)
ds1 <- subset(ds1, select = c(SPCOND, pH, ORP, AlkalinityW, NH3_f, H2S_f, NO3_f, Cl, Fl, NO2, NO3, SO4, Li, Na_IC, K_IC, Mg_IC, Ca_IC, Al, Si, P, Ca, Cr, Mn, Fe, Cu, Zn, Sr, K, Mg, Na, O18, D)) # Pulling out some variables
nb <- estim_ncpPCA(ds1, method.cv = "Kfold")
ggplot(data = NULL, aes(x = 0:5, y = nb$criterion)) + 
    geom_point() + 
    geom_line() + 
    xlab("nb dim") + 
    ylab("MSEP")
```
We can see that the MSEP is minimized when S is 2, which we can access with `nb$ncp`. Let's perform the actual PCA imputation.

```{r Imputing DS2018}
res.imp <- imputePCA(ds1, ncp = 2, scale = TRUE) # from previous block
res.pca <- PCA(res.imp$completeObs) # uses PCA function from FactoMineR package
```

Let's do the same for US2018.

```{r US2018 PCA}
## HM.....let's just take out the variables that matter, for example in DS2018
us1 <- data.frame(US2018) # Take a copy of the data

# let's also set ID as the rowname
rownames(us1) <- us1$ID

# Let's try looking at all good enough variables. First, I'll use a function to help determine the value of dimension S that that minimizes the mean squared error of prediction (MSEP)
us1 <- subset(us1, select = c(SPCOND, pH, ORP, AlkalinityW, NH3_f, H2S_f, NO3_f, Cl, Fl, NO2, NO3, SO4, Li, Na_IC, K_IC, Mg_IC, Ca_IC, Al, Si, P, Ca, Cr, Mn, Fe, Cu, Zn, Sr, K, Mg, Na, O18, D)) # Pulling out some variables
nb <- estim_ncpPCA(us1, method.cv = "Kfold")
ggplot(data = NULL, aes(x = 0:5, y = nb$criterion)) + 
    geom_point() + 
    geom_line() + 
    xlab("nb dim") + 
    ylab("MSEP")

res.imp <- imputePCA(us1, ncp = nb$ncp, scale = TRUE) # from previous block
res.pca <- PCA(res.imp$completeObs) # uses PCA function from FactoMineR package
```

What if I did multple imputations? Using the MIPCA function.

```{r MIPCA DS}
resMIPCA_ds <- MIPCA(ds1, ncp = 2, nboot = 200)
plot(resMIPCA_ds)
```

```{r MIPCA US}
resMIPCA_us <- MIPCA(us1, ncp = 2, nboot = 200)
plot(resMIPCA_us)
```

What if I only pulled out what was conservative? These include $\delta^{18}$O, $\delta$D, Cl$^-$, and SO$_4$$^{2-}$, but might be safe to include our base cations and maybe alkalinity and SPCOND. Let's give it a try!

```{r Conservative MIPCA for DS2018}
ds2 <- data.frame(DS2018)
rownames(ds2) <- ds2$ID

us2 <- data.frame(US2018)
rownames(us2) <- us2$ID

ds2 <- subset(ds2, select = c(SPCOND, AlkalinityW, Cl, SO4, Na_IC, K_IC, Mg_IC, Ca_IC, O18, D))
us2 <- subset(us2, select = c(SPCOND, AlkalinityW, Cl, SO4, Na_IC, K_IC, Mg_IC, Ca_IC, O18, D))

# first find out what the minimal S is
nb_ds <- estim_ncpPCA(ds2, method.cv = "Kfold")
nb_us <- estim_ncpPCA(us2, method.cv = "Kfold")

# Multiple imputation
resMIPCA_ds <- MIPCA(ds2, ncp = 3, nboot = 200)
plot(resMIPCA_ds)
```

```{r Conservative MIPCA for US2018}
resMIPCA_us <- MIPCA(us2, ncp = 3, nboot = 200)
plot(resMIPCA_us)
```

OK, now what if we just considered the REAL conservative solutes? We don't really have to imputate then, but it'd be interesting to do anyways.

```{r Conservative Conservative DS2018}
ds3 <- data.frame(DS2018)
rownames(ds3) <- ds3$ID

us3 <- data.frame(US2018)
rownames(us3) <- us3$ID

ds3 <- subset(ds3, select = c(Cl, SO4, O18, D))
us3 <- subset(us3, select = c(Cl, SO4, O18, D))

dsres.pca <- PCA(ds3)
```

```{r Conservative Conservative US2018}
usres.pca <- PCA(us3)
```

Hm....so maybe I didn't need to impute at all? Maybe I could impute the endmembers. Checking to see if I get similar results using `ggfortify`.

```{r Reproducibility}
library(ggfortify)
pca_res_ds <- prcomp(ds3, scale. = TRUE)
autoplot(pca_res_ds, label = TRUE, label.size = 3, shape = FALSE, loadings = TRUE,
         loadings.label = TRUE, loadings.label.size = 5)
```

Out of curiosity, let's check out a few more things.

```{r Scree plot}
# To extract eigenvalues/variances
library(factoextra)
get_eig(dsres.pca)
# Visualize eigenvalues/variances
fviz_screeplot(dsres.pca, addlabels = TRUE)
```

Extract and visualize results for variables.
```{r More vis}
var <- get_pca_var(dsres.pca)
var
head(var$coord)
head(var$contrib)
fviz_pca_var(dsres.pca, col.var = "black")
```

```{r}
# Control variable colors using their contributions
fviz_pca_var(dsres.pca, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping
             )
```

Contribution of variables to the principal axes:

```{r Variable contribution}
# Contributions of variables to PC1
fviz_contrib(dsres.pca, choice = "var", axes = 1, top = 10)

# Contributions of variables to PC2
fviz_contrib(dsres.pca, choice = "var", axes = 2, top = 10)
```

Extracting results for individuals

```{r Individuals}
ind <- get_pca_ind(dsres.pca)
fviz_pca_ind(dsres.pca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )
```

Biplot of individuals and variables
```{r Biplot}
fviz_pca_biplot(dsres.pca, repel = TRUE)
```
Total contribution of individuals to PC 1-2
```{r Contribution of individuals to S = 2}
fviz_contrib(dsres.pca, choice = "ind", axes = 1:2)
```


OK now let's add the supplementary info for individuals (end members). For this go, I'm going to toss every possible endmember on there, except for the US samples.

```{r}
soil1 <- data.frame(SOIL2018)
rownames(soil1) <- soil1$ID
soil3 <- subset(soil1, select = c(Cl, SO4, O18, D))

spring1 <- data.frame(SPRING2018)
# let's just get rid of "sp box sample" for now
spring1 <- spring1[-c(1),] # remove the first row, now n = 2
rownames(spring1) <- spring1$ID
spring3 <- subset(spring1, select = c(Cl, SO4, O18, D))

rain1 <- data.frame(RAIN2018)
rownames(rain1) <- rain1$ID
rain3 <- subset(rain1, select = c(Cl, SO4, O18, D))


# time to merge vertically
# library(tidyverse)
em <- bind_rows(ds3,soil3,spring3,rain3)

supplres.pca <- PCA(em, ind.sup = 22:31)
```

More on individuals:
```{r Predicted results for the supplementary individuals}
supplres.pca$ind.sup
```
Visualizing supplementary individuals

```{r Supplementary visualization}
p <- fviz_pca_ind(supplres.pca, col.ind.sup = "blue", repel = TRUE)
p
```

OK, let's see what a crazy mess this will look like with the upstream samples added as supplementary individuals.

```{r Upstream as Individuals}
em2 <- bind_rows(ds3,us3,soil3,spring3,rain3)
supplres2.pca <- PCA(em2, ind.sup = 22:56)
```

Ok that's gross, but there's a couple things I can tease out. I can have four endmembers: Rain (average of Rain 1 and Rain 2), average of soil seeps (Well 203 and Well 205), Well 200 and HW Trib (bowl shaped, probably has older soil water, most likely closest to actual groundwater; HW Trib could be related to Well 200), and maybe the average of the upstream samples? Maybe use median instead.

OK, let's take the median of the US points and means for the Rain and Well (W3 and W5) samples.

```{r Taking median/mean values for endmembers}
library(matrixStats)
library(plyr)

us3_median <- matrixStats::colMedians(as.matrix(us3))
us3_std <- matrixStats::colSds(as.matrix(us3))

rain3_mean <- matrixStats::colMeans2(as.matrix(rain3))
rain3_std <- matrixStats::colSds(as.matrix(rain3))

# ok dealing with the soil samples
## first extracting W0 and HW Trib
deepsoil3_mean <- matrixStats::colMeans2(as.matrix(soil3[c(1,3),])) # row 1 = HW Trib, row 3 = Well 200 / W0 
deepsoil3_std <- matrixStats::colSds(as.matrix(soil3[c(1,3),]))

topsoil3_mean <- matrixStats::colMeans2(as.matrix(soil3[c(5,6),])) # row 5 = Well 203 / W3, row 6 = Well 205 / W5
topsoil3_std <- matrixStats::colSds(as.matrix(soil3[c(5,6),]))

# Convert vectors into dataframe in 1 row
us3_median <- matrix(us3_median, ncol = 4)
us3_median <- as.data.frame(us3_median)
rownames(us3_median) <- "US"

rain3_mean <- matrix(rain3_mean, ncol = 4)
rain3_mean <- as.data.frame(rain3_mean)
rownames(rain3_mean) <- "Rain"

deepsoil3_mean <- matrix(deepsoil3_mean, ncol = 4)
deepsoil3_mean <- as.data.frame(deepsoil3_mean)
rownames(deepsoil3_mean) <- "Deep Soil"

topsoil3_mean <- matrix(topsoil3_mean, ncol = 4)
topsoil3_mean <- as.data.frame(topsoil3_mean)
rownames(topsoil3_mean) <- "Shallow Soil"

# Ok let's merge these endmember dataframes and then rename the columns to match ds3. Let's also ignore the error/stdev for now.
endmembers <- bind_rows(us3_median,rain3_mean,deepsoil3_mean,topsoil3_mean)
endmembers <- rename(endmembers, c("V1"= "Cl", "V2" = "SO4", "V3" = "O18", "V4" = "D"))

# Finally let's join this with DS3 as supplementary individuals
emmas <- bind_rows(ds3,endmembers)
emmas.pca <- PCA(emmas, ind.sup = 22:25)
```

OK, so the US endmember is pretty bad, but let's just keep it moving? Luckily, by the Rule of One, I could just reduce the components to U1 and U2 and I would need three endmembers: Deep Soil, Top Soil, Rain, which beautifully bound the stream samples except for DS22 and DS20. 

Let's have a PCA where US wasn't included, which would greatly reduce the problem to three-endmember mixing.

```{r Without US}
endmembers2 <- bind_rows(rain3_mean,deepsoil3_mean,topsoil3_mean)
endmembers2 <- rename(endmembers2, c("V1"= "Cl", "V2" = "SO4", "V3" = "O18", "V4" = "D"))

emmas2 <- bind_rows(ds3,endmembers2)
emmas2.pca <- PCA(emmas2, ind.sup = 22:24)
```

```{r}
dsc <- rgb(0.06,0.22,0.67) # blue
xyz <- data.frame(emmas2.pca$ind.sup$coord)
xyz <- subset(xyz, select = c(Dim.1,Dim.2))
# fviz_pca_ind(emmas2.pca, repel = TRUE)
# plot.PCA(emmas2.pca)
## S3 method for class 'PCA'
indie <- plot(emmas2.pca, axes = c(1, 2), choix = c("ind","var","varcor"),
    ellipse = NULL, xlim = NULL, ylim = NULL, habillage="none", 
    col.hab = NULL, col.ind=dsc, col.ind.sup="black", 
    col.quali="magenta", col.quanti.sup="blue", col.var="black",
    label = c("ind.sup", "ind"),
	invisible = c("none","ind","ind.sup","quali","var","quanti.sup"), 
    lim.cos2.var = 0., title = NULL, palette=NULL,
    autoLab = c("auto","yes","no"), new.plot = FALSE, select = c("DS17","DS11","DS22","DS20","Deep Soil","Shallow Soil","Rain"), 
	unselect = 0, shadowtext = FALSE, legend = list(bty = "y", x = "topleft"),
	graph.type = c("ggplot","classic"), ggoptions = list(size = 7))

indie_final <- indie + theme(panel.grid = element_blank(),
              axis.title = element_text(size = 24),
              axis.text = element_text(size = 24, color = "black"),
              plot.title = element_text(size = 24),
              legend.position = "none") +
  geom_polygon(data = xyz, aes(x = Dim.1, y = Dim.2, alpha = 0.1), colour = "grey")
# default
# plot(x, axes = c(1, 2), choix = c("ind","var","varcor"),
#     ellipse = NULL, xlim = NULL, ylim = NULL, habillage="none", 
#     col.hab = NULL, col.ind="black", col.ind.sup="blue", 
#     col.quali="magenta", col.quanti.sup="blue", col.var="black",
#     label = c("all","none","ind","ind.sup","quali","var","quanti.sup"),
# 	invisible = c("none","ind","ind.sup","quali","var","quanti.sup"), 
#     lim.cos2.var = 0., title = NULL, palette=NULL,
#     autoLab = c("auto","yes","no"), new.plot = FALSE, select = NULL, 
# 	unselect = 0.7, shadowtext = FALSE, legend = list(bty = "y", x = "topleft"),
# 	graph.type = c("ggplot","classic"), ggoptions = NULL, ...)

ggsave(file = "indie.pdf", plot = indie_final, width = 10.25, height = 6, dpi = 600)
 
```


```{r GSA Individuals Plot}
indie2 <- plot(emmas2.pca, axes = c(1, 2), choix = c("ind","var","varcor"),
    ellipse = NULL, xlim = NULL, ylim = NULL, habillage="none", 
    col.hab = NULL, col.ind=dsc, col.ind.sup="black", 
    col.quali="magenta", col.quanti.sup="blue", col.var="black",
    label = c("none"),
	invisible = c("none","ind","ind.sup","quali","var","quanti.sup"), 
    lim.cos2.var = 0., title = NULL, palette=NULL,
    autoLab = c("no"), new.plot = FALSE, select = NULL, 
	unselect = 0, shadowtext = FALSE, legend = list(bty = "y", x = "topleft"),
	graph.type = c("ggplot","classic"), ggoptions = list(size = 7))

indie2_final <- indie2 + theme(panel.grid = element_blank(),
              axis.title = element_text(size = 22),
              axis.text = element_text(size = 20, color = "black"),
              plot.title = element_text(size = 22),
              legend.position = "none",
              panel.border = element_rect(colour = "black", fill=NA, size=2)) +
  geom_polygon(data = xyz, aes(x = Dim.1, y = Dim.2, alpha = 0.9))
# default
# plot(x, axes = c(1, 2), choix = c("ind","var","varcor"),
#     ellipse = NULL, xlim = NULL, ylim = NULL, habillage="none", 
#     col.hab = NULL, col.ind="black", col.ind.sup="blue", 
#     col.quali="magenta", col.quanti.sup="blue", col.var="black",
#     label = c("all","none","ind","ind.sup","quali","var","quanti.sup"),
# 	invisible = c("none","ind","ind.sup","quali","var","quanti.sup"), 
#     lim.cos2.var = 0., title = NULL, palette=NULL,
#     autoLab = c("auto","yes","no"), new.plot = FALSE, select = NULL, 
# 	unselect = 0.7, shadowtext = FALSE, legend = list(bty = "y", x = "topleft"),
# 	graph.type = c("ggplot","classic"), ggoptions = NULL, ...)

ggsave(file = "indie2.pdf", plot = indie2_final, width = 6.34, height = 4, dpi = 600)
 
```

```{r GSA Var}
indie2 <- plot(emmas2.pca, axes = c(1, 2), choix = c("var"),
    ellipse = NULL, xlim = NULL, ylim = NULL, habillage="none", 
    col.hab = NULL, col.ind="black", col.ind.sup="blue", 
    col.quali="magenta", col.quanti.sup="blue", col.var="black",
    label = c("none"),
	invisible = c("none","ind","ind.sup","quali","var","quanti.sup"), 
    lim.cos2.var = 0., title = NULL, palette=NULL,
    autoLab = c("auto","yes","no"), new.plot = FALSE, select = NULL, 
	unselect = 0.7, shadowtext = FALSE, legend = list(bty = "y", x = "topleft"),
	graph.type = c("ggplot","classic"), ggoptions = NULL)

indie2_final <- indie2 + theme(panel.grid = element_blank(),
              axis.title = element_text(size = 22),
              axis.text = element_text(size = 20, color = "black"),
              plot.title = element_text(size = 22),
              legend.position = "none",
              panel.border = element_rect(colour = "black", fill=NA, size=2)) 
# default
# plot(x, axes = c(1, 2), choix = c("ind","var","varcor"),
#     ellipse = NULL, xlim = NULL, ylim = NULL, habillage="none", 
#     col.hab = NULL, col.ind="black", col.ind.sup="blue", 
#     col.quali="magenta", col.quanti.sup="blue", col.var="black",
#     label = c("all","none","ind","ind.sup","quali","var","quanti.sup"),
# 	invisible = c("none","ind","ind.sup","quali","var","quanti.sup"), 
#     lim.cos2.var = 0., title = NULL, palette=NULL,
#     autoLab = c("auto","yes","no"), new.plot = FALSE, select = NULL, 
# 	unselect = 0.7, shadowtext = FALSE, legend = list(bty = "y", x = "topleft"),
# 	graph.type = c("ggplot","classic"), ggoptions = NULL, ...)

ggsave(file = "var.pdf", plot = indie2_final, width = 7.34, height = 4, dpi = 600)
 
```

OK now finally we can do EMMA. These calculations should be based on the following sets of equations from Christopherson et al. (1990) to solve the contribution of each end member to streamflow:

\[
1 = a_1 + a_2 + a_3
\]

\[
SW_{U1} = a_1EM_{1U1}+a_2EM_{2U1}+a_3EM_{3U1}
\]

\[
SW_{U2} = a_1EM_{1U2}+a_2EM_{2U2}+a_3EM_{3U2}
\]

where $a_1$, $a_2$m and $a_3$ are the fractions of each end member, $SW_{U1}$ and $SW_{U2}$ are the project stream water observations in U space coordinates, and $EM_{1U1}$ and $EM_{3U2}$ are the coefficients of the $n$th end member projected in U space.

Let's extract some data:
```{r See description of PCA output}
emmas2.pca
```
```{r}
em_coords <- emmas2.pca$ind.sup$coord
em_coords <- as.data.frame(em_coords)
dsu <- emmas2.pca$ind$coord
dsu <- as.data.frame(dsu)

# subsetting DS discharge
dsq <- subset(DS2018, select = c(SamplingDate, Q))

# Let's remove Howler 2
dsu <- dsu[-c(21),]
dsq <- dsq[-c(21),]
```

OK, let's solve these one by one.

```{r DS1: Flowpaths}
A = array(c(3.529428, -2.704799, 1, -2.656899, -1.596949, 1, 5.134543, 3.986612, 1
), dim = c(3,3)) # rain, deep soil, top soil 
b = c(0.08410838*0.003010143,-1.1266902*0.003010143, 0.003010143)
solve(A,b)
```

```{r DS1: Fractions}
A = array(c(3.529428, -2.704799, 1, -2.656899, -1.596949, 1, 5.134543, 3.986612, 1), dim = c(3,3)) # rain, deep soil, top soil 
b = c(0.08410838,-1.1266902, 1)
solve(A,b)
```

Yay!! I have real numbers for DS1. OK, now figuring out how to automate this process. I have three dataframes to pull stuff from: em_coords (end member projected U-space coordinates), dsu (DS U-space coordinates), dsq (stream hydrograph data).

```{r Fractional Contributions}
# OK very sleep deprived and may make some of this kinda lazy
## First, I will assign variables for EM1U1, EM1U2, EM2U1, EM2U2, EM3U1, EM3U2 (which are rain, deep, and top)
r1 <- em_coords[1,1] # ex. Rain 1 in U1
r2 <- em_coords[1,2]
d1 <- em_coords[2,1]
d2 <- em_coords[2,2]
t1 <- em_coords[3,1]
t2 <- em_coords[3,2]

datalist = data.frame()
A <- array(c(r1, r2, 1, d1, d2, 1, t1, t2, 1), dim = c(3,3))

for (i in 1:nrow(dsu)) {
  b <- c(dsu$Dim.1[i],dsu$Dim.2[i],1)
  data <- array(solve(A,b), dim = c(1,3)) 
  datafr = data.frame(data)
  datalist <- rbind(datalist, datafr)
}

```

```{r Flowpaths}
# The only thing that is different is multiply b by the Q at DS[i]
r1 <- em_coords[1,1]
r2 <- em_coords[1,2]
d1 <- em_coords[2,1]
d2 <- em_coords[2,2]
t1 <- em_coords[3,1]
t2 <- em_coords[3,2]

flowpath = data.frame()
A <- array(c(r1, r2, 1, d1, d2, 1, t1, t2, 1), dim = c(3,3))

for (i in 1:nrow(dsu)) {
  b <- c(dsu$Dim.1[i]*dsq$Q[i],dsu$Dim.2[i]*dsq$Q[i],1*dsq$Q[i])
  data <- array(solve(A,b), dim = c(1,3)) 
  datafr = data.frame(data)
  flowpath <- rbind(flowpath, datafr)
}
```

```{r Concatenating}
# First let's concatenate the fractions and fractional discharges with DS Time and Q
datalist <- datalist %>% as.data.frame() %>% dplyr::rename(f_rain = X1,
                           f_deep = X2,
                           f_top  = X3)

flowpath <- flowpath %>% as.data.frame() %>%
  dplyr::rename(Q_rain = X1,
    Q_deep = X2,
    Q_top  = X3)

dsemma <- cbind(dsq,datalist,flowpath)
```

FINALLY! Let's export datalist (fractions) and dsemma (contributions) and make the figures.

```{r Exporting as csv}
write.csv(dsemma,"dsemma.csv")
```


```{r}
#~~~~~~~~~~~~~~ Discharge ~~~~~~~~~~~~~~#
DownQ <- read.csv("https://raw.githubusercontent.com/wdnguyen/CR/master/lia_Q.csv")
DownQ$Date <- as.POSIXct(DownQ$Date, format ="%m/%d/%y %H:%M")

ggplot() +
  geom_area(data = DownQ, aes(x = Date, y = Q, colour="Total Q")) +
  geom_area(data = dsemma, aes(x = SamplingDate, y = Q_deep, fill="Q_deep")) +
  geom_area(data = dsemma, aes(x = SamplingDate, y = Q_top, fill= "Q_top")) +
  geom_area(data = dsemma, aes(x = SamplingDate, y = Q_rain, fill= "Q_rain")) +
  scale_color_manual("",
                     breaks = c("Total Q", "Q_rain", "Q_deep","Q_top"),
                     values = c("Total Q"="black", "Q_rain" = "blue", "Q_deep" = "red","Q_top" = "green")) +
  scale_fill_manual("",
                    breaks = c("Total Q", "Q_rain", "Q_deep","Q_top"),
                    values = c("Total Q"="black", "Q_rain" = "blue", "Q_deep" = "red","Q_top" = "green")) +
  theme_bw()
```

```{r Time series without filling}
#~~~~~~~~~~~~~~ Discharge ~~~~~~~~~~~~~~#
ggplot() +
  geom_area(data = DownQ, aes(x = Date, y = Q, colour="Total Q")) +
  geom_area(data = dsemma, aes(x = SamplingDate, y = Q_deep, colour="Q_deep", fill = "red")) +
  geom_area(data = dsemma, aes(x = SamplingDate, y = Q_top, colour= "Q_top", fill = "green")) +
  geom_area(data = dsemma, aes(x = SamplingDate, y = Q_rain, colour= "Q_rain", fill = "blue")) +
  scale_color_manual("",
                     breaks = c("Total Q", "Q_rain", "Q_deep","Q_top"),
                     values = c("Total Q"="black", "Q_rain" = "blue", "Q_deep" = "red","Q_top" = "green")) +
  scale_fill_manual("",
                    breaks = c("Total Q", "Q_rain", "Q_deep","Q_top"),
                    values = c("Total Q"="black", "Q_rain" = "blue", "Q_deep" = "red","Q_top" = "green")) +
  theme_bw()
```
```{r Time series of fractions over time}
library(lubridate)
library(scales)

lims <- as.POSIXct(strptime(c("2018-07-06 12:00", "2018-07-09 00:00"), 
                   format = "%Y-%m-%d %H:%M"))

# https://stackoverflow.com/questions/59229001/how-to-set-breaks-with-exact-interval-in-scale-x-datetime-in-ggplot
t_breaks = seq(from = ymd_hms('2018-07-06 12:00:00'), to = ymd_hms('2018-07-09 00:00:00'), by = '12 hours')

frac <- ggplot(data = dsemma, aes(x = SamplingDate)) +
    annotate(geom = "rect", alpha = .1, ymin = 0, ymax = 1, xmin = as.POSIXct('07/06/18 12:00',format ="%m/%d/%y %H:%M"),
    xmax = as.POSIXct('07/09/18 00:00', format ="%m/%d/%y %H:%M")) +
  geom_line(aes(y = f_rain, colour = "Rain")) +
  geom_line(aes(y = f_deep, colour = "Deep Soil")) +
  geom_line(aes(y = f_top, colour = "Shallow Soil")) +
  geom_point(aes(y = f_rain, colour = "Rain")) +
  geom_point(aes(y = f_deep, colour = "Deep Soil")) +
  geom_point(aes(y = f_top, colour = "Shallow Soil")) +
  scale_color_manual("",
                     breaks = c("Rain", "Deep Soil","Shallow Soil"),
                     values = c("Rain" = "blue", "Deep Soil" = "red","Shallow Soil" = "green")) +
  theme_few() +
  scale_x_datetime(labels = date_format("%m/%d\n%H:%M", tz = "GMT"), 
                        limits = lims, 
                   breaks = t_breaks, # date_breaks("12 hour"), # as.POSIXct(c("2018-07-07 00:00","2018-07-07 12:00", "2018-07-08 00:00", "2018-07-08 12:00", "2018-07-09 00:00")), 
                        expand = c(0, 0)) +
  theme(axis.text = element_text(size = 24, family = 'sans'),
        axis.text.x = element_text(size = 24, family = 'sans', colour = 'black'),
        axis.text.y = element_text(size = 24, family = 'sans', colour = 'black'),
        axis.title.y = element_text(size = 24, family = 'sans'),
        axis.title.x = element_text(size = 24, family = 'sans'),
        legend.text = element_text(size = 24, family = 'sans'),
        legend.title = element_blank(),
        plot.margin = unit(c(2,1,1,1), "cm")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Date", y = "Contribution")

# limits=c(as.POSIXct("2018-07-06 12:00:00",tz="UTC"), as.POSIXct("2018-07-09 00:00:00",tz="UTC")) in scale_x_datetime

ggsave(file = "frac.pdf", plot = frac, width = 10.25, height = 6, dpi = 600)
```

```{r}
rain <- read.csv("https://raw.githubusercontent.com/wdnguyen/CR/master/DownstreamMiller.csv")
rain$Date <- as.POSIXct(rain$Date, format ="%m/%d/%y %H:%M")

# Rain
rain_plot <- ggplot(data = rain[1:581,], aes(x = Date, y = Rain)) + 
  geom_bar(stat = "identity", 
           position = 'dodge',
           # fill = "black", #rgb(187, 8, 38, maxColorValue = 255), 
           color = "black") + #rgb(187, 8, 38, maxColorValue = 255)) + 
  theme_bw() + 
  # ggtitle("48-Hour Data Slam") +
  theme(#text=element_text(size=18),
        panel.border = element_blank(), 
        panel.grid.major.x = element_blank(), 
        panel.grid.major.y = element_blank(),# element_line(colour = "black"), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"), 
        axis.text.x = element_blank(), #(angle = 45, hjust = 1, color="white"),
        axis.line.x = element_blank(), #(color = "white"),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(colour="black", size=18),
        axis.title.x = element_blank(),
        axis.title.y = element_text(colour="black", size=18)) + 
  labs(y = "Precip\n(mm)") +
  # scale_y_reverse() + 
  scale_y_continuous(trans = "reverse")
```

```{r PHREEQC}
library(reshape2)
library(tidyverse)
library(viridis)

phr <- read.csv("phreeqc_1.csv")
names(phr)[1] <- "ID"
phr_melt <- melt(data = phr, id.vars = "ID")

phreeqc <- ggplot(data = phr_melt, aes(x = variable, y = value, fill = ID)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_fill_viridis(discrete = T, option = "E") +
  ggtitle("Possible Phase Mole Transfers from US to DS") +
  labs(x = "Phase", y = "Phase Mole Transfers (mol)") +
  theme_bw() +
  theme(axis.text = element_text(size = 18, color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(size = 18, color = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE))

ggsave(file = "phreeqc.pdf", plot = phreeqc, width = 7, height = 4.5, dpi = 600)
```

